Client-Side Indexes for Fast Full-Text Searching

### 2014-12-11 paper/paper.tex
fix auther section
clean up intro and related work

\numberofauthors{3}

\author{
\alignauthor
Amy X. Zhang\\
       \email{axz@mit.edu}
\alignauthor Lea Verou\\
       \email{leaverou@mit.edu}
\alignauthor Manali Naik\\
       \email{manalinaik@mit.edu}
}

@@

Many applications on the web use a combination of client-side and server-side data stores to facilitate fast interactive and data-intensive experiences. 
However, standard client side databases within browsers do not currently support full-text searching. 
In this paper, we describe a client-side search engine built on top of IndexedDB that makes use of several types of indexes common to many well-known server-side search engines.
We compare the performance of different indexes on different types of full-text content and queries and find that....
We also compare the performance of our system with that of fully server side systems and examine scenarios where a hybrid approach may be fastest. We find that...

\end{abstract}

@@

Most web applications store the bulk of their data on a backend web server or on the cloud, and the client side queries the server when it needs to access data.
As web applications have become more interactive and responsive in the past several years, developers have built larger and more complex web servers to support them. However, this separation of data from the client introduces significant network effects that can often lead to noticeable latency, as data may be retrieved from servers that are very far away.

In recent years, more applications are being built that heavily rely on client-side storage and offload data processing to web browsers. Today most commonly used browsers support the current HTML5 standards, which include a persistent client-side DBMS with the name IndexedDB that is accessible via a set of JavaScript APIs. 
The ability to use local storage enables real-time interactivity on data-intensive and collaborative tasks by reducing the number of network roundtrips.
These have been useful for many large and complex web applications such as massive multi-player games CITE that would have previously only existed as desktop applications.
Issues around privacy and ownership of data have also given rise to a number of applications CITE that move most or all of data storage to the client~\cite{bilenko2011predictive}. 
Finally, many popular smartphone browsers rely on wireless signals to transmit information, which is prone to failure in areas with poor signal, suggesting that applications that are entirely client-side or applications that retains some functionality during downtime could be useful~\cite{balasubramanian2012findall}.

While much work has been done to make it easier to store and query data in the browser~\cite{benson2010sync}, currently there are few known ways to conduct full-text searching on the client side. Search engines are a very important component of a large number of applications and are necessary to make sense of many forms of full-text data such as chat or private messages, social media content, search queries, and webpage titles and content. 
As personalization has become more important for search engines, many designers of systems and architectures have highlighted the opportunities for client-side data computation~\cite{bharat2000searchpad,teevan2005personalizing}.
Issues of privacy are also important when it comes to social applications and searching over personal data. New social applications such as CITE have been built that can operate without any servers, instead using peer-to-peer or Bluetooth connections.
However, while it is possible to store full-text content on the client side, there is still no way to conduct full-text search on the client side at a reasonable speed, preventing most data computation necessary for a fully-featured search engine.
The current standard of IndexedDB, a key-value store, only supports simple operations such as put and get on set of keys and secondary indexes on object fields. There are no built-in capabilities to create indexes for full-text content.

In this paper, we introduce Lucy.js, a fully client-side JavaScript search engine that is an extension of the current IndexedDB API CITE. It has been designed to be easy for anyone currently using IndexedDB in their application to begin indexing and searching full-text content, without needing to modify pre-existing code or move to a new framework or mode of communication. There are also no dependencies on a server or on a network, so it can be harnessed offline. In our implementation, we created several types of indexes commonly used in well-known search engines, such as Lucene CITE and within PostgreSQL CITE, including inverted indexes, prefix and suffix tries, and lossy search trees. We discuss how we managed to create these indexes on top of the IndexedDB framework as well as how the different indexes are suited to different tasks. We compare their performance on different types of queries and with different sizes and types of datasets. We find that...

Finally, we imagine the potential real-world use cases for a client-side search engine and how performance may be improved using this system. We implement a set of applications that would be reasonable in the real world and measure performance between fully client-side, fully server-side, and mixtures of both databases to respond to queries. We find that....
This research demonstrates not only the feasibility of having search engines within the browser but also the potential for new applications and improved performance of existing applications through the additional capabilities provided by a client-side search engine.

@@

Though previous HTML standards and browser capabilities were more limited, the idea of using of the client side for data storage and data computation has been around for many years. 
Some research has looked into building tools to support an architecture that includes client-side databases, including toolkits to synchronize data between the client and server~\cite{benson2010sync}. Another project facilitates ubiquitous access to client-side data across multiple browsers by enabling browser session migration from client to server to client~\cite{lo2013imagen}.

When it comes to search engines, specifically web search, prior work has found that 40\% to 60\% of search queries are re-finding queries~\cite{teevan2005personalizing}. This finding has lead to applications that store web content from the result of a search engine to improve speed and availability when the network is down~\cite{bozzon2010liquid,bharat2000searchpad,balasubramanian2012findall}. Prior work in this area have also highlighted how web search can often be a highly personal activity, and that web search should be personalized to the user to improve accuracy. Personal information can then be stored on the client side as this information is primarily useful for this particular client~\cite{teevan2005personalizing}.
Privacy is also a large concern for applications that collect personal data. Some applications have tackled personalized advertising using client-side stored profiles to mitigate this problem~\cite{bilenko2011predictive}.

With the increased use of mobile phones in recent years, many web applications are now accessed via a smartphone. Here too, studies have shown that many queries to search engines are for re-finding. Using a client-side database is potentially even more impactful for a mobile phone browser because network connectivity is much less robust. Researchers have demonstrated this feasiblity by building a standalone search application for Android~\cite{balasubramanian2012findall}. However, this search engine is not usable with other browsing apps or other phones. Because we build an extension on top of IndexedDB, our search engine will be able to work for all browsers, desktop or mobile, that support IndexedDB, which is the majority of them.

Some prototypes of client-side search engines have previously been built CITE. Currently, all existing implementations only create basic inverted indexes.
For instance, previous research has demonstrated that it is feasible to store an inverted index within IndexedDB~\cite{lin:jscene} though it is much slower than using a server-side application such as Lucene.
We improve on these works by implementing different types of indexes for full-text search and comparing their performance. 
Another existing implementation of a client side search engine called Lunr.js\footnote{http://www.lunrjs.com} stores indexes in the client-side memory, which is infeasible for large amounts of data and also non-persistent. Instead, our system uses IndexedDB object stores to hold indexes. Also, the YDN-DB module\footnote{https://github.com/yathit/ydn-db-fulltext} which wraps around the IndexedDB API, requires developers to migrate over to use the YDN system. Instead our system is simply an extension of IndexedDB.
We also demonstrate several real-life situations that may call for a hybrid client and server database architecture for full-text search and measure their speed.

@@

>>>>

\numberofauthors{1}

\author{
\alignauthor Amy X. Zhang, Lea Verou, Manali Naik\\
\vspace{.2cm}
	\affaddr{MIT CSAIL}\\
	\affaddr{Cambridge, MA USA} \\
       \email{\{axz, leaverou, mnaik\}@mit.edu}
}

@@

Many applications on the web today use a combination of client-side and server-side data stores to facilitate fast interactive and data-intensive experiences. 
However, standard client side databases within browsers do not currently support full-text searching, and it is not clear the best way to engineer full-text indexes on top of the existing infrastructure.
In this paper, we introduce Lucy.js, a fully client-side search engine built as an extension of the existing well-supported IndexedDB API.
We compare the performance of the different indexes we implement on different types of full-text content and queries. We find that a hybrid approach combining inverted indexes and tries perform the best, particularly on larger datasets.
We also implement various real-life applications of our system and compare the performance of our system with that of using a fully server side system. We find that given the costs of network latency, our approach is signficantly faster than relying on a server-side search engine.
Finally, we discuss how our system may open up novel web application and data storage designs that solely use client-side search engines or combine both client and server search engines.
\end{abstract}

@@

A typical web application stores the bulk of its data on a backend web server or on the cloud, and the client side queries the server when it needs to access data.
As web applications have become more interactive and responsive in the past several years, developers have built larger and more complex web servers to support them. However, this separation of data from the client introduces significant network effects that can often lead to noticeable latency, as large amounts of data may need to be retrieved from servers that are very far away.

In recent years, more applications have been built that heavily rely on client-side storage and offload data processing to web browsers. Today most commonly used browsers support the current HTML5 standards, which include a persistent client-side DBMS with the name IndexedDB that is accessible via a set of JavaScript APIs. 
The ability to use local storage enables real-time interactivity on data-intensive and collaborative tasks by reducing the number of network roundtrips.
These have been useful for many large and complex web applications such as massive multi-player games and collaborative editing tools that would have previously only existed as desktop applications.
Issues around privacy and ownership of data have also given rise to a number of applications that aim to move sensitive data analysis and data storage to the client~\cite{bilenko2011predictive, khoshgozaran2009private}. 
Finally, most smartphone browsers rely on wireless signals to transmit information, which is prone to failure in areas with poor signal, suggesting that applications that are entirely client-side or applications that retains some functionality during downtime could be useful~\cite{balasubramanian2012findall}.

While much work has focused on making client-side storage and query easier~\cite{benson2010sync}, currently there are few known ways to conduct full-text searching on the client side. This is a problem because search engines are a very important component of a large number of applications and are necessary to make sense of many forms of full-text data such as chat or private messages, social media content, search queries, and webpage titles and content. 
As personalization has become more important for search engines, many designers of systems and architectures have highlighted the opportunities for client-side data computation~\cite{bharat2000searchpad,teevan2005personalizing}.
Issues of privacy are also important when it comes to social applications and searching over personal data.
However, while it is possible to store full-text content on the client side, there is still no way to conduct full-text search on the client side at a reasonable speed, preventing most data computation necessary for a fully-featured search engine.
The current standard of IndexedDB, a key-value store, only supports simple operations such as put and get on set of keys and secondary indexes on object fields. There are no built-in capabilities to create indexes for full-text content.

In this paper, we introduce Lucy.js, a fully client-side JavaScript search engine that is an extension of the current IndexedDB API~\cite{indexeddb}. It has been designed to be easy for anyone currently using IndexedDB in their application to begin indexing and searching full-text content, without needing to modify pre-existing code or move to a new framework or mode of communication. There are also no dependencies on a server or network, so it can be harnessed offline. In our implementation, we created several types of indexes commonly used in well-known search engines, such as Lucene and within PostgreSQL, including inverted indexes, prefix and suffix tries, and variations on these. We discuss how we create these indexes on top of the IndexedDB framework as well as how the different indexes are suited to different tasks. We compare their performance on different types of queries and with different sizes and types of datasets. We find that...FILL IN HERE.

We evaluate our system on several real-world applications and compare performance of fully client-side and fully server-side data architectures. We find that on the whole, using client-side indexes significantly decreases the time to complete search queries due to the network latency brought on by communicating with a remote server.
Finally, we consider potential novel systems and architectures that could be implemented that take advantage of a client-side search engine.
This research demonstrates not only the feasibility of having fast search engines within the browser but also the improved performance of existing applications and the potential for new applications through the additional capabilities provided by a client-side search engine.

@@

Though previous HTML standards and browser capabilities were more limited, the idea of using of the client side for data storage and data computation has been around for many years~\cite{bharat2000searchpad}. 
Some research has looked into building tools to support an architecture that includes client-side databases, including toolkits to synchronize data between the client and server~\cite{benson2010sync}. Another project facilitates ubiquitous access to client-side data across multiple browsers by enabling browser session migration from client to server to client~\cite{lo2013imagen}.

When it comes to search engines, specifically web search, prior work has found that 40\% to 60\% of search queries are re-finding queries~\cite{teevan2005personalizing}. This finding has led to applications that store web content from the result of a search engine to improve speed and availability when the network is down~\cite{bozzon2010liquid,bharat2000searchpad,balasubramanian2012findall}. Prior work in this area have also highlighted how web search can often be a highly personal activity, and that web search could be personalized to the user to improve accuracy. Personal information can then be stored on the client side as this information is primarily useful for this particular client~\cite{teevan2005personalizing}.
Privacy is also a large concern for applications that collect personal data. Some applications have tackled personalized advertising using client-side stored profiles to mitigate this problem~\cite{bilenko2011predictive}.

With the increased use of mobile phones in recent years, many web applications are now accessed via a smartphone. Here too, studies have shown that many queries to search engines are for re-finding. Using a client-side database is potentially even more impactful for a mobile phone browser because network connectivity is much less robust. Researchers have demonstrated this feasiblity by building a standalone search application for Android~\cite{balasubramanian2012findall}. However, this search engine is not usable with other browsing apps or other phones. Because we build an extension on top of IndexedDB, our search engine will be able to work for all browsers, desktop or mobile, that support IndexedDB.

Some prototypes of client-side search engines have previously been built. However, all existing implementations only create basic inverted indexes.
For instance, previous research has demonstrated that it is feasible to store an inverted index within IndexedDB~\cite{lin:jscene} though it was found to be much slower than using a server-side application such as Lucene.
We improve on these works by implementing different types of indexes for full-text search and comparing their performance. 
Another existing implementation of a client side search engine called Lunr.js\footnote{http://www.lunrjs.com} stores indexes in the client-side memory, which is infeasible for large amounts of data and also non-persistent. Instead, our system uses IndexedDB object stores to hold indexes. Also, the YDN-DB module\footnote{https://github.com/yathit/ydn-db-fulltext} which implements an inverted index, wraps around the IndexedDB API, requiring developers to migrate over to use the YDN system. Instead our system is simply an extension of the IndexedDB API, making it easy to integrate.
We also demonstrate several real-life situations that may call for a hybrid client and server database architecture for full-text search and measure their speed.

@@

Newer peer-to-peer standards such as WebRTC for browsers may make more client-heavy social and interactive applications feasible in the future.
